---
title: "Alignment-based workflow"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---



First set up your Zorn/Bascet workdirectory as before. If you wish to run
these steps on a SLURM cluster, see separate vignette and adapt accordingly.

```{r}
library(Zorn)
bascet_runner <- LocalRunner(direct = TRUE, show_script=TRUE)
bascetRoot = "/home/yours/an_empty_workdirectory"
```

We will use BWA for alignment, which wants the data in FASTQ format. Here we assume that you use a single-cell
WGS chemistry that produces paired reads, and direct Zorn to produce R1/R2 FASTQ files (only R1 specified).
We name the output Bascets "asfq", taking as input the typical sharded reads:

(SLURM-compatible)

```{r}
### Get reads in fastq format
BascetMapTransform(
  bascetRoot,
  "filtered",   #default; can omit
  "asfq",  ###name parameter
  out_format="R1.fq.gz",
  runner=bascet_runner
)
```



Before alignment, you need to prepare your reference by indexing it.
You can do this using "bwa index" in the terminal; but you can also do
it via Zorn. This assumes that you have a FASTA-file named all.fa
(can also be gzipped):

(SLURM-compatible)

```{r}
#Build a reference
BascetIndexGenomeBWA(
  bascetRoot, 
  "/your_reference/all.fa", 
  runner=bascet_runner
)
```





You can now proceed with alignment using BWA. By default, this command
outputs both unsorted and sorted BAM-files.

* The sorted BAM-files are required for counting of reads in genomic regions, i.e., WGS species counting and RNA-seq
* The unsorted BAM-files are required for filtering of host reads

(SLURM-compatible)

```{r}
### Perform alignment
BascetAlignToReference(
  bascetRoot,
  useReference="/your_reference/all.fa",
  numLocalThreads=10,
  runner=bascet_runner
)
```


TODO: Filtering

==================== Host filtering workflow ==============



```{r}
BascetFilterAlignment(
  bascetRoot,
  outputName="filter_paired",
  keep_mapped=TRUE,
  runner=bascet_runner
)
```









==================== Counting workflow ==============


If you want to analyze where reads mapped, either for RNA-seq counting for genes,
or more generally, what the count is within a genomic region, you can use Signac
for great flexibility in this task. Signac is designed to analyze
single-cell ATAC-seq data, where the reads are stored in a specialized BED-file.
You can use the command below to generate this file.

Note one caveat: depending on chemistry, this workflow may not handle UMIs,
and just report raw read counts. However, for quick and dirty operations
this is typically enough.

(SLURM-compatible)

```{r}
### Generate fragments BED file suited for quantifying reads/chromosome using Signac later 
BascetBam2Fragments(
  bascetRoot,
  runner=bascet_runner
)
```



If you just want to get the number of reads per chromosome (i.e. to compare a metagenomic sample to a number of defined strains),
then we provide a simpler counter. The output is a small matrix that you can load with Seurat:

(SLURM-compatible)

```{r}
### Count reads per chromosome
BascetCountChrom(
  bascetRoot, 
  runner=bascet_runner
)
```


