---
title: "De novo assembly"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


First set up Zorn/Bascet according to the [install instructions](install.html). This tutorial assumes that you have [debarcoded the reads](debarcoding.html).

Assembly of each cell's genome can then be performed using SKESA:

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}
### Assemble all genomes
BascetMapCellSKESA(
  bascetRoot,
  inputName = "filtered",
  outputName = "contigs"
)
```

This produce contigs for each cell, all grouped together in "contigs". Note that
both Skesa and Spades only assembles cells for which there are sufficient number of reads.


# Investigating assemblies

The output is zip-files with contigs for each cell, separately. You can thus easily extract the contigs of
interest and analyze with any tool of choice (not just Zorn/Bascet).

You can obtain the contigs for one cell like this (this is the fastest way):
```{r, eval=F, echo=T}
b <- OpenBascet(bascetRoot, "contigs", bascetInstance)
contigs <- BascetReadFile(b, cellID = "<cell_id>", filename = "contigs.fa", as = "text")
CloseBascet(b)
```


We also provide convencience functions for getting a subset. The worked example below will work for future file formats as well, not just ZIP:

```{r, eval=F, echo=T}
bascetRoot <- "/path/to/your/bascet"
bascetInstance <- GetDefaultBascetInstance()

# Get cell names
cells <- BascetCellNames(bascetRoot, "contigs", bascetInstance = bascetInstance)

# Pick the cells you want — e.g. all of them, or a subset
listCells <- cells$cell[1:10]

# Make sure the output dir exists
outputDir <- "contigs_out"
dir.create(outputDir, showWarnings = FALSE)

# Dump contigs — one <cellid>.fa file per cell
BascetDumpContigs(
  bascetRoot     = bascetRoot,
  inputName      = "contigs",          # shard holding contigs.fa per cell
  listCells      = listCells,
  outputDir      = outputDir,
  bascetInstance = bascetInstance
)
```


# Assembly analysis

To assess assemblies, have a look at our [Map scripts](map_scripts.html) vignette, which shows how to run per-cell tools
(e.g. QUAST for assembly QC).







# Coassembly

########################################################### TODO ###########################

If you have cells with highly similar genomes, you might be able to generate
"coassemblies" - consensus assemblies using reads from multiple similar cells.

To do this, simply extract the debarcoded reads from the cells of interest.
There are many ways of picking them, where one method is to use the clustering
function in Seurat. But you can use any method you wish to come up with a
list of cell names

First set up your Zorn/Bascet work directory as before.

Then this is a very random method to extract cells after clustering ([see this Seurat tutorial first](https://satijalab.org/seurat/articles/pbmc3k_tutorial.html)).
It is an open research problem how to best carry this out:

```{r, eval=F, echo=T}
### Get cell names
cell_names <- rownames(adata[adata$cluster_id="0",])
```

Next extract a FASTQ with reads:

```{r, eval=F, echo=T}
### Extract reads TODO
Bascet....A(
  bascetRoot,
  inputName = "filtered",
  outputName = "contigs"
)
```

And finally, run [SKESA](https://github.com/ncbi/SKESA) or your favourite software to assemble the contigs: (in BASH)


```{r, eval=F, echo=T}

#SKESA etc

```



