---
title: "De novo assembly"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---



First set up your Zorn/Bascet work directory as before. Assembly of each cell's genome can then be performed using SKESA:

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}
### Assemble all genomes
BascetMapCellSKESA(
  bascetRoot,
  inputName = "filtered",
  outputName = "contigs"
)
```

This produce contigs for each cell, all grouped together in "contigs". Note that
both Skesa and Spades only assembles cells for which there are sufficient number of reads.


# Investigating assemblies

The output is zip-files with contigs for each cell, separately. You can thus easily extract the contigs of
interest and analyze with any tool of choice (not just Zorn/Bascet).

You can obtain the contigs for one cell like this (this is the fastest way):
```{r, eval=F, echo=T}
b <- OpenBascet(bascetRoot, "contigs", bascetInstance)
contigs <- BascetReadFile(b, cellID = "<cell_id>", filename = "contigs.fa", as = "text")
CloseBascet(b)
```


We also provide convencience functions for getting a subset. The worked example below will work for future file formats as well, not just ZIP:

```{r, eval=F, echo=T}
bascetRoot <- "/path/to/your/bascet"
bascetInstance <- GetDefaultBascetInstance()

# Get cell names
cells <- BascetCellNames(bascetRoot, "contigs", bascetInstance = bascetInstance)

# Pick the cells you want — e.g. all of them, or a subset
listCells <- cells$cell[1:10]

# Make sure the output dir exists
outputDir <- "contigs_out"
dir.create(outputDir, showWarnings = FALSE)

# Dump contigs — one <cellid>.fa file per cell
BascetDumpContigs(
  bascetRoot     = bascetRoot,
  inputName      = "contigs",          # shard holding contigs.fa per cell
  listCells      = listCells,
  outputDir      = outputDir,
  bascetInstance = bascetInstance
)
```


# Assembly analysis

To assess assemblies, have a look at our [Map scripts](map_scripts.html) vignette, which shows how to run per-cell tools
(e.g. QUAST for assembly QC).
