---
title: "Working with Bascet-ZIP files"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---

If you want to get files out of Bascets, this the tutorial for you! Note that
you should avoid doing this unless needed - the map system is there if you wish
to process each cell; aggregate can then be used to systematically extract information.
These offer the benefit of multithreading and you do not need to worry about having
copious amounts of files which definitely will not make your systems administrator happy.

That said, in case of emergency, or if you just want a single file, this is the
approach to take.

First load an instance, and ensure that it works:
```{r, eval=F, echo=T}
bascetInstance <- getBascetSingularityImage(...)
TestBascetInstance(bascetInstance)
```

You can now open a file. Internally, this creates an instance of Bascet running
in a separate thread, which R can send commands to. This design is necessary to
avoid the startup cost of Docker and Singularity, such that you pay the cost only
once; but reading many files is very fast.
```{r, eval=F, echo=T}
skesa_file <- OpenBascet(
  bascetRoot, 
  "skesa", 
  bascetInstance
)
```


You can obtain a list of files for a cell like this (see reference for more commands):
```{r, eval=F, echo=T}
BascetListFilesForCell(
  skesa_file, 
  "G1_E5_H7_E11", 
  bascetInstance = bascetInstance
)
```


A common scenario is that you want the assembled genome of a particular cell. It
can be done like this:
```{r, eval=F, echo=T}
contig_file <- BascetReadFile(
  skesa_file,
  cellID = "G1_E5_H7_E11", 
  filename = "contigs.fa",
  as="text"
)
```


Be sure to close the file when you are done, as you have a rather large ecosystem
of tools still in memory!
```{r, eval=F, echo=T}
CloseBascet(
  skesa_file
)
```

You can now save the file to disk. If you want to actually look at the data,
we recommend loading it using the seqinr package rather than directly using the
raw file data obtained by the function above.
```{r, eval=F, echo=T}
writeLines(
  contig_file, 
  "my_genome.fa"
)
```



