---
title: "Coassembly"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


If you have cells with highly similar genomes, you might be able to generate
"coassemblies" - consensus assemblies using reads from multiple similar cells.

To do this, simply extract the debarcoded reads from the cells of interest.
There are many ways of picking them, where one method is to use the clustering
function in Seurat. But you can use any method you wish to come up with a
list of cell names

First set up your Zorn/Bascet work directory as before.

Then this is a very random method to extract cells after clustering ([see this Seurat tutorial first](https://satijalab.org/seurat/articles/pbmc3k_tutorial.html)).
It is an open research problem how to best carry this out:

```{r, eval=F, echo=T}
### Get cell names
cell_names <- rownames(adata[adata$cluster_id="0",])
```

Next extract a FASTQ with reads:

```{r, eval=F, echo=T}
### Extract reads TODO
Bascet....A(
  bascetRoot,
  inputName = "filtered",
  outputName = "contigs"
)
```

And finally, run [SKESA](https://github.com/ncbi/SKESA) or your favourite software to assemble the contigs: (in BASH)


```{r, eval=F, echo=T}

#SKESA etc

```




