---
title: "Doublet removal"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


First set up Zorn/Bascet according to the [install instructions](install.html). This tutorial assumes that you have already produced a Seurat object using one of the count workflows, e.g. the [KRAKEN2 workflow](kraken.html), the [alignment-based workflow](alignment.html), the [informative KMER workflow](kmer.html), or the [count sketch workflow](countsketch.html).
You should have a count table of some sorts ready for analysis.

# When to remove doublets

A *doublet* in single-cell metagenomics is a barcode that captures DNA from more than one cell.
These can arise for at least two common reasons:

1. If more cells are collected than there are barcodes, you get **barcode collisions**. Droplet and split-pool methods rely on the cells being much fewer than the avaiable barcodes. By doing "Poisson loading" (diluting enough), doublets are typically kept <2%; but they still occur and must be handled

2. If cells stick together, the will also receive the same barcode. Ensuring a good single-cell suspension is crucial

Zorn/Bascet produces count tables and you can thus use many single-cell tools to process the data, such as to remove doublets.
Here we give an example using **`scDblFinder`**.

# Removing doublets using scDblFinder

[scDblFinder](https://bioconductor.org/packages/release/bioc/html/scDblFinder.html) lives in Bioconductor and operates on a `SingleCellExperiment` object,
so the workflow is to wrap the count matrix from your Seurat object, run the
caller, then copy the per-cell scores back into the Seurat object's metadata.

```{r, eval=F, echo=T}
library(SingleCellExperiment)
library(scDblFinder)

## Wrap the raw counts as a SingleCellExperiment (rows = features, cols = cells)
sce <- SingleCellExperiment::SingleCellExperiment(
  list(counts = GetAssayData(adata,layer = "counts"))
)

## Call doublets
sce <- scDblFinder::scDblFinder(sce)

## Copy the per-cell score and class back into the Seurat object
adata$scDblFinder.score <- sce$scDblFinder.score
adata$scDblFinder.class <- sce$scDblFinder.class
```

You can now overlay the score and class on the UMAP to see whether doublets
form a coherent population (often they sit between two species clusters, or they all form a "blob" in the middle of the UMAP):

```{r, eval=F, echo=T}
FeaturePlot(adata, features = "scDblFinder.score")
DimPlot(adata, group.by = "scDblFinder.class")
```

To remove doublets, just subset the Seurat object:

```{r, eval=F, echo=T}
adata <- adata[, adata$scDblFinder.class == "singlet"]
```
