---
title: "Genome annotation tools"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


# QUAST


First run QUAST on all the cells:
[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCell(
  bascetRoot,
  withfunction = "_quast",
  inputName = "skesa",  #or other source of contigs
  outputName = "quast"
)
```


Then aggregate the results for visualization.
This example caches the result to speed up reloading; this is optional
```{r, eval=F, echo=T}
quast_aggr <- BascetCacheComputation(bascetRoot,"cache_quast",MapListAsDataFrame(BascetAggregateMap(
  bascetRoot,
  "quast",
  aggr.quast
)))
```




# Abricate


First run Abricate on all the cells. The NCBI database is used by default.
See ListDatabaseAbricate() for a list of other databases

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellAbricate(
  bascetRoot,
  inputName = "skesa"  #or other source of contigs
)
```



Then aggregate the results for visualization.
This example caches the result to speed up reloading; this is optional
```{r, eval=F, echo=T}
quast_aggr <- BascetCacheComputation(bascetRoot,"cache_abricate",MapListAsDataFrame(BascetAggregateMap(
  bascetRoot,
  "abricate",
  aggr.abricate
)))
```



# Bakta


First download a database:
```{r, eval=F, echo=T}
DownloadDatabaseBakta(
  dbdir="~/bakta",  #create directory before running command
  dbtype="light"
)
```

You can run then run Bakta on all cells:
[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
### ....
BascetMapCellBakta(
  bascetRoot,
  db = "~/bakta",
  inputName = "skesa",  #or other source of contigs
)
```





Then aggregate the results for visualization.
This example caches the result to speed up reloading; this is optional
```{r, eval=F, echo=T}
quast_aggr <- BascetCacheComputation(bascetRoot,"cache_bakta",MapListAsDataFrame(BascetAggregateMap(
  bascetRoot,
  "bakta",
  aggr.bakta
)))
```



