---
title: "Genome annotation tools"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


This vignette covers the per-cell genome annotation wrappers. All of them
are built on the MAP framework — see [Map scripts](map_scripts.html) for the
underlying mechanism and how to write your own.


# QUAST — assembly quality

QUAST computes standard assembly QC metrics from contigs (N50, number and
length of contigs, GC content, mis-assemblies, etc.). It is the most common
way to compare the quality of different assemblies and is the de-facto
standard for reporting assembly statistics.

* Website: <https://github.com/ablab/quast>
* Cite: Gurevich A, Saveliev V, Vyahhi N, Tesler G. *QUAST: quality assessment tool for genome assemblies.* Bioinformatics. 2013;29(8):1072–1075.

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellQUAST(
  bascetRoot,
  inputName  = "contigs",  #or other source of contigs
  outputName = "quast"
)
```

Then aggregate the results for visualization.
This example caches the result to speed up reloading; this is optional
```{r, eval=F, echo=T}
quast_aggr <- BascetCacheComputation(bascetRoot,"cache_quast",MapListAsDataFrame(BascetAggregateMap(
  bascetRoot,
  "quast",
  aggr.quast
)))
```


# Abricate — AMR / virulence screening

Abricate performs mass screening of contigs against several curated
databases of antimicrobial resistance and virulence genes (e.g. NCBI, CARD,
ResFinder, VFDB, PlasmidFinder). It only reports acquired resistance
genes, not point mutations, and is widely used because it is fast and easy
to interpret.

* Website: <https://github.com/tseemann/abricate>
* Cite: Seemann T. *Abricate*. GitHub https://github.com/tseemann/abricate (please also cite the underlying database you used)

The NCBI database is used by default. See `ListDatabaseAbricate()` for a
list of other databases.

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellAbricate(
  bascetRoot,
  inputName  = "contigs",  #or other source of contigs
  outputName = "abricate",
  db         = "ncbi"
)

abricate_mat <- BascetAggregateAbricate(
  bascetRoot,
  inputName = "abricate"
)
```


# Bakta — genome annotation

Bakta is a fast, standardized annotation tool for bacterial genomes and
plasmids. It identifies coding sequences, ncRNAs, tRNAs, CRISPR arrays,
and more, and assigns functional descriptions through alignment-free
sequence identification. It is a good choice when you want a complete
genome annotation comparable across cells.

* Website: <https://github.com/oschwengers/bakta>
* Cite: Schwengers O, Jelonek L, Dieckmann MA, Beyvers S, Blom J, Goesmann A. *Bakta: rapid and standardized annotation of bacterial genomes via alignment-free sequence identification.* Microbial Genomics. 2021;7(11):000685.

First download a database:

```{r, eval=F, echo=T}
DownloadDatabaseBakta(
  dbdir  = "~/bakta",  #create directory before running command
  dbtype = "light"
)
```

You can then run Bakta on all cells:

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellBakta(
  bascetRoot,
  inputName  = "contigs",  #or other source of contigs
  outputName = "bakta",
  db         = "~/bakta"
)
```

Then aggregate the results for visualization.
This example caches the result to speed up reloading; this is optional
```{r, eval=F, echo=T}
bakta_aggr <- BascetCacheComputation(bascetRoot,"cache_bakta",MapListAsDataFrame(BascetAggregateMap(
  bascetRoot,
  "bakta",
  aggr.bakta
)))
```


# Ariba — AMR identification from reads

Ariba detects antimicrobial resistance genes (and other gene panels)
directly from sequencing reads, without first assembling. It builds local
assemblies around reference genes and reports SNPs and indels relative to
the reference, which makes it useful when assembly quality is too low to
trust contig-based screening.

* Website: <https://github.com/sanger-pathogens/ariba>
* Cite: Hunt M, Mather AE, Sánchez-Busó L, Page AJ, Parkhill J, Keane JA, Harris SR. *ARIBA: rapid antimicrobial resistance genotyping directly from sequencing reads.* Microbial Genomics. 2017;3(10):e000131.

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellAriba(
  bascetRoot,
  inputName  = "filtered",
  outputName = "ariba",
  db         = "/path/to/ariba_db/out.prepareref"
)

ariba_mat <- BascetAggregateAriba(
  bascetRoot,
  inputName = "ariba"
)
```


# AMRfinder — NCBI AMRfinderPlus

AMRfinderPlus screens contigs (or proteins) against the NCBI Reference
Gene Catalog to identify acquired AMR genes, point mutations conferring
resistance, and selected virulence and stress-response genes. It is
maintained by NCBI and is the source for many downstream AMR databases.

* Website: <https://github.com/ncbi/amr>
* Cite: Feldgarden M, Brover V, Gonzalez-Escalona N, Frye JG, Haendiges J, Haft DH, Hoffmann M, Pettengill JB, Prasad AB, Tillman GE, Tyson GH, Klimke W. *AMRFinderPlus and the Reference Gene Catalog facilitate examination of the genomic links among antimicrobial resistance, stress response, and virulence.* Scientific Reports. 2021;11(1):12728.

First download the database:

```{r, eval=F, echo=T}
DownloadDatabaseAMRfinder("/path/to/amrfinder_db")
```

Then run on all cells:

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellAMRfinder(
  bascetRoot,
  inputName  = "contigs",
  outputName = "AMRfinder",
  db         = "/path/to/amrfinder_db"
)

amr_df <- BascetAggregateAMRfinder(
  bascetRoot,
  inputName = "AMRfinder"
)
```


# GECCO — biosynthetic gene clusters

GECCO predicts biosynthetic gene clusters (BGCs) in assembled contigs
using a conditional random field over Pfam domain compositions. It is
much faster than antiSMASH and is well suited to scanning thousands of
single-cell assemblies for natural-product potential.

* Website: <https://gecco.embl.de/>
* Cite: Carroll LM, Larralde M, Fleck JS, Ponnudurai R, Milanese A, Cappio Barazzone E, Zeller G. *Accurate de novo identification of biosynthetic gene clusters with GECCO.* bioRxiv. 2021. doi:10.1101/2021.05.03.442509

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellGECCO(
  bascetRoot,
  inputName  = "contigs",
  outputName = "gecco"
)
```

Aggregate the per-cell cluster tables into a single list of data.frames:

```{r, eval=F, echo=T}
gecco_aggr <- BascetAggregateGECCO(
  bascetRoot,
  inputName = "gecco"
)
```
