---
title: "Installation"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


First install Zorn. It is under active development, so just get it straight
from GitHub and install from source:

```{r, eval=F, echo=T}
if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
devtools::install_github("henriksson-lab/zorn")
```

You also need Bascet. Unless you are a Bascet developer, you will find it
easiest to run it through Singularity or Docker. Pick the one that is easiest
to set up on your particular system (i.e. if you already have one of these set
up, go with that option)

# Accessing Bascet via Singularity

First [install Singularity](https://docs.sylabs.io/guides/3.0/user-guide/installation.html)

Once installed, Zorn can then pull down the latest Singularity image for you:

```{r, eval=F, echo=T}
bascet_inst <- getBascetImageInstance(store_at="/somewhere/on/your/disk/")  ## todo docker
```


# Accessing Bascet via Docker

First [install Docker](https://docs.docker.com/get-started/)

Once installed, Zorn can then pull down the latest Docker image for you:

```{r, eval=F, echo=T}
bascet_inst <- getBascetImageInstance(store_at="/somewhere/on/your/disk/")  ## todo docker
```


# Using the Bascet instance

You need to pass the bascet_inst variable to each Zorn command that requires it.
The singularity image will be cached, so if you run this again, it will instead use
the previous image. You will need to delete it manually if you wish to replace it.


# For developers: Accessing a locally compiled Bascet

See [separate page](for_developers.html)

