---
title: "Installation"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---

## Step 1: Install Zorn

Zorn is under active development, so just get it straight
from GitHub and install from source:

```{r, eval=F, echo=T}
if (!requireNamespace("devtools", quietly = TRUE))
    install.packages("devtools")
devtools::install_github("henriksson-lab/zorn")
```

You can then load it:
```{r, eval=F, echo=T}
library(Zorn)
```

## Step 2: Install Bascet

You also need Bascet, our command-line suite for single-cell analysis. Most users will find it easiest to run Bascet through Singularity or Docker; instructions for each are described below. Singularity is for Linux users, while OSX and Windows users have to use Docker.

### Install Bascet via Singularity (Linux only)

First [install Singularity](https://docs.sylabs.io/guides/3.0/user-guide/installation.html).

Note one possibly annoying default setting! We could not read Bascet files on some file systems, because they were invisible to Bascet.
To avoid this, edit /etc/singularity/singularity.conf to below:
```{}
mount hostfs = yes
```

Once installed, Zorn can then pull down the latest Singularity image for you.
We recommend keeping it in a separate directory from your workspace, as
multiple workspaces can share the same image

```{r, eval=F, echo=T}
bascet_instance.default <- getBascetSingularityImage(store_at="/somewhere/on/your/disk/")
```


### Install Bascet via Docker (Singularity is recommended for Linux)

First [install Docker](https://docs.docker.com/get-started/).

Once installed, Zorn can then pull down the latest Docker image for you.
We recommend keeping it in a separate directory from your workspace, as
multiple workspaces can share the same image

```{r, eval=F, echo=T}
bascet_instance.default <- getBascetDockerImage(store_at="/somewhere/on/your/disk/")
```

## Step 3: Test the installation

Run the following and see if you get "ok" before proceeding:
```{r, eval=F, echo=T}
TestBascetInstance(bascet_instance.default)
```

## Using the Bascet instance

You need to pass the bascet_inst variable to each Zorn command that requires it.
The singularity image will be cached, so if you run this again, it will instead use
the previous image. You will need to delete it manually if you wish to replace it.

## For developers: Accessing a locally compiled Bascet

See [Bascet for Developers](for_developers.html)

