---
title: "KRAKEN2 "
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


First set up your Zorn/Bascet working directory as before. If you wish to run
these steps on a SLURM cluster, see separate vignette and adapt accordingly.

```{r, eval=F, echo=T}
library(Zorn)
bascetRunner.default <- LocalRunner(direct = TRUE, showScript=TRUE)
bascetRoot <- "/home/yours/an_empty_workdirectory"
```

# Classifying reads with KRAKEN2

To run KRAKEN2, you need a database. You can get them here:
https://benlangmead.github.io/aws-indexes/k2

If you want a small one then consider standard-8. Unzip it in a directory. You
can then run KRAKEN2 like this:

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}
### Run Kraken on each cell. Produce a count matrix of taxonomy features
BascetRunKraken(
  bascetRoot,
  useKrakenDB="/your_disk/kraken/standard-8"
)
```

Internally, there are two steps here. First KRAKEN2 taxonomically classifies each read.
In the second step, we count the taxonomic reads for each cell.
This ends up being a rather small count matrix that you can process using Seurat.


# Postprocessing with Signac/Seurat

The output format is one binary HDF5 file for each shard, roughly in the
[Anndata file format](https://anndata.readthedocs.io/en/stable/fileformat-prose.html).
To load these, use the following command that loads the files and concatenates them
into a single matrix. It can then be loaded into Seurat:

```{r, eval=F, echo=T}
mat <- ReadBascetCountMatrix(bascetRoot,"kraken", verbose=FALSE)

taxid_ob <- CreateAssayObject(t(mat)) #t() needed to conform from anndata to seurat
adata <- CreateSeuratObject(counts = taxid_ob, project = "proj", min.cells = 0, min.features = 0) 
```

You then most likely want to know which cell is which species. Zorn can annotate species
by simply picking the dominant species, phylum and group for each cell:

```{r, eval=F, echo=T}
## Look up taxonomy consensus data given taxonomyID counts for each cell
kraken_taxid <- KrakenFindConsensusTaxonomy(mat)

## Add KRAKEN consensus taxonomy to metadata
rownames(kraken_taxid) <- kraken_taxid$cell_id
kraken_taxid <- kraken_taxid[colnames(adata),c("taxid","phylum","class","order","family","genus","species")] #optional subsetting
adata@meta.data <- cbind(
  adata@meta.data,
  kraken_taxid[colnames(adata),c("taxid","phylum","class","order","family","genus","species")]
)
```


You will need to filter low-abundance cells. To do this, first investigate
a kneeplot of each species. As different species are differently hard to lyse,
you will likely have phylum-specific kneeplot patterns.

```{r, eval=F, echo=T}
showNumSpecies <- 10
KrakenKneePlot(adata, groupby = "phylum", showNumSpecies=showNumSpecies)
```


You can then proceed to filter out cells with few reads. Note that a general cutoff
is nonideal as different phyla have different amount of DNA
(due to different genome sizes & lysis biases); the best way to handle
this is an open research question (e.g. you could also consider different cutoffs
for different species). This is however the most basic filtering you can perform:

```{r, eval=F, echo=T}
keep_cells <- adata$nCount_RNA > 10000 #10k reads
sum(keep_cells)
adata <- adata[,keep_cells]
```


You can now proceed to generate a dimensional reduction in the form of a UMAP.
See the tutorials for [Seurat](https://satijalab.org/seurat/) and
[Signac](https://stuartlab.org/signac/) for details. We summarize the required
command below, where you can apply either an RNA-seq or ATAC-seq dimensional reduction
depending on what you think is appropriate.

```{r, eval=F, echo=T}
if(TRUE){
  #ATAC-seq style analysis
  library(Signac)
  adata <- RunTFIDF(adata)
  adata <- FindTopFeatures(adata, min.cutoff = 'q0')
  adata <- RunSVD(adata)
  DepthCor(adata)
  adata <- RunUMAP(object = adata, reduction = 'lsi', dims = 1:30, reduction.name = "kraken_umap")  
} else {
  #RNA-seq style analysis
  adata <- NormalizeData(adata)
  adata <- FindVariableFeatures(adata, selection.method = "vst", nfeatures = 2000)
  adata <- ScaleData(adata, features = rownames(adata))
  adata <- RunPCA(adata, features = VariableFeatures(object = adata))
  adata <- RunUMAP(adata, dims = 1:20, reduction.name = "kraken_umap")
}
```

You can now plot a UMAP with your cells! Here colored by genus, but you can also try other levels of annotation:

```{r, eval=F, echo=T}
DimPlot(object = adata, label = TRUE, group.by = "genus", reduction = "kraken_umap") + 
  NoLegend() + 
  xlab("KRAKEN1") + 
  ylab("KRAKEN2")
```



