---
title: "Downloading NCBI reference genomes"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---

# When to use this

Zorn/Bascet can download public NCBI assemblies and pack them directly into
Bascet-ZIP shards. Each assembly is treated as one cell, and each cell contains
one file

# Download the metadata directory

This downloads the RefSeq bacterial assembly summary from NCBI and caches it
under the user's cache directory:

```{r, eval=FALSE}
metadataFile <- BascetDownloadNcbiGenomeMetadata(
  db = "refseq",
  group = "bacteria"
)

assemblies <- BascetReadNcbiGenomeMetadata(metadataFile)
dim(assemblies)
```

# Pick genomes

The table can be treated like a data.frame in R, and filtered to get the genomes
you are interested in. This is just a random example of filtering:

```{r, eval=FALSE}
selected <- assemblies[
  assemblies$version_status == "latest" &
    (is.na(assemblies$excluded_from_refseq) | assemblies$excluded_from_refseq == "") &
    assemblies$refseq_category %in% c("reference genome", "representative genome") &
    !is.na(assemblies$ftp_path) &
    nzchar(assemblies$ftp_path),
]
```

# Download NCBI genomes

We will run several downloads in parallel, each creating one shard. To do this,
we first pepare input lists for each shard:

```{r, eval=FALSE}
BascetPrepareNcbiGenomeFetchLists(
  bascetRoot = bascetRoot,
  selected
)
```

Then start the download jobs:

```{r, eval=FALSE}
BascetDownloadNcbiGenomes(
  bascetRoot = bascetRoot
)
```

By default, Bascet writes ZIP shards. To write TIRP shards instead, set the
suggested output extension:

```{r, eval=FALSE}
BascetDownloadNcbiGenomes(
  bascetRoot = bascetRoot,
  inputName = "tofetch",
  outputName = "filtered",
  outFormat = "tirp.gz"
)
```

# NCBI load vs SLURM

The Bascet downloader uses a global download-start rate limit inside each job.
The default in the Zorn wrapper is conservative to avoid problems.
Note that if you launch many array jobs at once (SLURM), the total rate is
multiplied by the number of simultaneously running jobs. If this is a problem
depends on things like the IP of the compute nodes etc.
