---
title: "Read-based quality control"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---


# FASTQC

(move to another section?)

If you want to run QC on all cells as a whole, to get the average picture,
simply run FASTQC on reads after transformation to FASTQ:

```{r, eval=F, echo=T}
### Get reads in fastq format
BascetMapTransform(
  bascetRoot,
  inputName="filtered",   
  outputName="asfq",      
  out_format="R1.fq.gz"
)
```

You can also run FASTQC on each individual cell, in which case you do not
need to convert to FASTQ as above. This takes a fair bit of time,
but can help tell if, e.g., a cluster of cells is caused by technical issues
such as adapter content. You first run FASTQC with mapcell:

[(SLURM-compatible step)](slurm.html)
```{r, eval=F, echo=T}
BascetMapCellFASTQ(
  bascetRoot,
  inputName = "filtered"  #or other source of reads
)
```


```{r, eval=F, echo=T}
BascetMapCellFASTQ(
  bascetRoot,
  inputName = "filtered"  #or other source of reads
)
```

If you have an outlier cell in your dataset, you can investigate its FASTQ
HTML report in the follow manner (opening in the RShiny plot pane, or separate browser):

```{r, eval=F, echo=T}
ShowFASTQCforCell(
    bascetFile, 
    cellID="xyz", #name of your cell 
    readnum="1", #for R1
)
```

You can also compare cells by aggregating the data. Note that FASTQC creates
rather complex statistics that need further extraction for simple plotting

```{r, eval=F, echo=T}
aggr_fastqc <- BascetAggregateFASTQC(
  bascetRoot
)
```


One relevant statistic is the adapter content across the read:

```{r, eval=F, echo=T}
PlotFASTQCadapterContent <- function(
    aggr_fastqc,
    readnum="1" #for R1
)
```


You can also retrieve a table of pass/fail statistics:

```{r, eval=F, echo=T}
fastqc_passfail <- GetFASTQCpassfailStats(
    aggr_fastqc,
    readnum="1" #for R1
)
```

Because there are so many things you can do with this statistics, we provide
a general interface to each table that FASTQC generates:
```{r, eval=F, echo=T}
mystats <- GetFASTQCassembledDF(
    aggr_fastqc, 
    section="see below", 
    readnum="1"
)
```

Possible values of section are:

* "Basic Statistics"
* "Per base sequence quality"
* "Per sequence quality scores"
* "Per base sequence content"
* "Per sequence GC content"
* "Per base N content"
* "Sequence Length Distribution"
* "Sequence Duplication Levels"
* "Overrepresented sequences"
* "Adapter Content"



