---
title: "SNP analysis"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---






# SNP-calling



Generate FASTQ-files for alignment

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}


###TODO

```



Align your reads using STAR (not BWA):

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}


###TODO



```






# Detect known SNPs

This workflow is useful if you wish to perform SNP-based donor deconvolution
using [Vireo](https://github.com/single-cell-genetics/vireo).
The [Vireo workflow](https://vireosnp.readthedocs.io/en/latest/genotype.html)
recommends obtaining a [pre-processed SNP list](https://sourceforge.net/projects/cellsnp/files/SNPlist/).

Given a VCF-file, containing your curated SNPs, to can run the following
command on your aligned reads (internally using cellsnp-lite):

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}
#TODO r command
#TODO option to not use UMI

#pip install -U vireoSNP

cellsnp-lite -s $BAM -b $BARCODE -O $OUT_DIR -R $REGION_VCF -p 10 --minMAF 0.1 --minCOUNT 20 --gzip

# https://cellsnp-lite.readthedocs.io/en/latest/
```


# Detect unknown SNPs




```{r, eval=F, echo=T}
#TODO r command
#TODO option to not use UMI
cellsnp-lite -s $BAM -b $BARCODE -O $OUT_DIR -p 10 --minMAF 0.1 --minCOUNT 100 --gzip
```



# Performing SNP-based donor deconvolution

This workflow relies on [Vireo](https://github.com/single-cell-genetics/vireo).

TODO option to provide donor genotype file



# https://vireosnp.readthedocs.io/en/latest/manual.html


# Citing

In addition to citing Zorn/Bascet, please also cite the following tools that
it wraps:

Xianjie Huang, Yuanhua Huang, Cellsnp-lite: an efficient tool for genotyping single cells, Bioinformatics, Volume 37, Issue 23, December 2021, Pages 4569–4571, https://doi.org/10.1093/bioinformatics/btab358




