---
title: "TCR/BCR analysis"
output:
  html_document:
    theme: united
    df_print: kable
  pdf_document: default
---





# Install TRUST4

TRUST4 is part of the Bascet container ([source](https://github.com/liulab-dfci/TRUST4)).
However, you need to provide a 



# The TCR/BCR workflow



Generate FASTQ-files for alignment

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}


###TODO

```



Align your reads using STAR (not BWA):

[(SLURM-compatible step)](slurm.html)

```{r, eval=F, echo=T}


###TODO

```



Run TRUST4 on your aligned reads:

[(SLURM-compatible step)](slurm.html)


```{r, eval=F, echo=T}
#TODO
run-trust4 -b foo.bam -f human_IMGT+C.fa --barcode BC --UMI UMI -t 10 
```


You can now load the extracted sequences in R:

```{r, eval=F, echo=T}
#TODO
```




# Citing

If you use TRUST4, you should cite it in addition to Zorn/Bascet
https://www.nature.com/articles/s41592-021-01142-2


Song, L., Cohen, D., Ouyang, Z. et al. TRUST4: immune repertoire reconstruction from bulk and single-cell RNA-seq data. Nat Methods (2021). https://doi.org/10.1038/s41592-021-01142-2





